---
title: "Data 607: Final Project"
author: "Farhana Akther"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction: 


```{r message=FALSE, warning=FALSE}
library(fs)
library(tidyverse)
library(DBI)
library(sparklyr)
library(zoo)
library(tidyquant)
library(plotly)
library(janitor)
library(timetk)
library(shiny)
library(shinythemes)
library(lubridate)
library(TTR)
```

# Part 1: Spark Installation

# Available for installation

```{r}
#spark_available_versions()
```

```{r}
#spark_installed_versions()
```


# Connects to spark Locally

```{r}
#?spark_connect
```

# configuration setup 

```{r}
# conf <- spark_config()
# conf$'sparklyr.cores.local' <- 4
# conf$'sparklyr.shell.driver-memory' <- "6g"
# conf$'spark.memory.fraction' <- 0.9
```

# Connect to spark Locally 1
```{r warning=FALSE, message=FALSE}
# config <- spark_config()
# config$spark.master <- "local[*]"
# sc<- spark_connect(config = config)
```

# Connect to spark Locally 2
```{r warning=FALSE, message=FALSE}
# sc<- spark_connect(
#   master = "local",
#   version = "3.3.2"
# )
```


# Disconnceting:
```{r}
#spark_disconnect_all(sc)
```

# web interface

```{r warning=FALSE, message=FALSE}
#spark_web(sc)
```


```{r}
#spark_disconnect_all(sc)
```

# PART 2:  FINANCIAL ANALYSIS---



## Loading Stock Data

```{r}
stocks_data_tbl <- read.csv("/Users/Admin/Desktop/FinalProject_Data607/fh_5yrs.csv")
```


```{r}
stocks_data_tbl$date<-as.Date(stocks_data_tbl$date) # changing dates from char to date format
```

```{r}
str(stocks_data_tbl)
```

# Filter Stock Data (ex: AAPL, Amzn etc)

```{r}
stocks <-stocks_data_tbl%>%
  filter(symbol %in% c('AAPL', 'AMZN', 'GOOG', 'MSFT', 'FB' ,'TSLA', 'CSCO', 'PEP', 'COST', 'SBUX', 'TEVA', 'INTC', 'GILD'))
```

# Copy Stocks example data to Spark

```{r}
# data_apple <-copy_to(
#   sc,
#   apple,
#   "apple_data",
#   Overwrite = TRUE
# )
```

How many rows (Spark)?


```{r}
#sdf_nrow(apple)
```


### Stockes Simple Moving Average (MOVING AVERAGE (SMA):

We will use *zoo* package to calculate a moving average (also called rolling average or running average) in R.



```{r warning=FALSE, message=FALSE}
# Convert data to zoo object
stocks.zoo <- zoo(stocks$adjclose, order.by = stocks$date)
```

```{r warning=FALSE}
# Calculate rolling statistics
stocksma <- rollmean(stocks.zoo, k = 20, na.pad = TRUE, align = "right")
stocksmax <- rollmax(stocks.zoo, k = 20, na.pad = TRUE, align = "right")
stocksmedian <- rollmedian(stocks.zoo, k = 20,  na.pad = TRUE, align = "right")
stockssum <- rollsum(stocks.zoo, k = 20, na.pad = TRUE, align = "right")
```


```{r}
# Combine data into one data frame
stocks_data <- data.frame(date = index(stocks.zoo),
                         adjclose = coredata(stocks.zoo),
                         ma = coredata(stocksma),
                         max = coredata(stocksmax),
                         median = coredata(stocksmedian),
                         sum = coredata(stockssum))
```

```{r}
stocks_data$date<-as.Date(stocks_data$date)
```

```{r}
str(stocks_data)
```

```{r, warning=FALSE}
# Create plot with multiple lines and a legend
ggplot(data = stocks_data, aes(x = date)) +
  geom_line(aes(y = adjclose, color = "Adjusted Close"), size = 0.8) +
  geom_line(aes(y = ma, color = "Moving Average"), size = 0.8) +
  geom_line(aes(y = max, color = "Moving Max"), size = 0.8) +
  geom_line(aes(y = median, color = "Moving Median"), size = 0.8) +
  geom_line(aes(y = sum, color = "Moving Sum"), size = 0.8) +
  scale_color_manual(values = c("blue", "red", "green", "purple", "orange"),
                     labels = c("Adjusted Close", "Moving Average", "Moving Max", "Moving Median", "Moving Sum")) +
  labs(x = "Date", y = "Price",
       title = "Stock Price and Rolling Statistics",
       subtitle = "2015-2020") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        legend.position = "left")
```

# SPARK + DPLYR: NASDAQ STOCK RETURNS ANALYSIS ---

```{r}
stocks_metrics_query <- stocks_data_tbl %>%

  group_by(symbol) %>%
  arrange(date) %>%
  mutate (lag = lag(adjclose, n = 1)) %>%
  ungroup() %>%

  mutate(returns = adjclose - lag) %>%
  
  group_by(symbol) %>%
  summarise(
   mean =  mean(returns, na.rm = TRUE),
   sd = sd(returns, na.rm = TRUE),
   count = n(),
   last_date = max(date, na.rm = TRUE)
  ) %>%
  ungroup()
```

```{r}
#stocks_metrics_query %>% show_query()
```

```{r}
#stocks_metric_tbl <- stocks_metrics_query %>% collect()
```

```{r}
#stocks_metric_tbl 
```

```{r}
#nasdaq_metric_tbl %>% write_rds("processed_data/nasdaq_metrics.rds")
```

## R DPLYR: APPLY SCREENING ---
- Metric Cap > $1B (More Stable)
- SD < 1 (Less  Volatile)
- Count > 3 * 365 (More stock history to base performance)
- Last Data = Max Date (Makes sure stock is still active)
- Reward Metric: variation of Sharpe Ratio (Mean Return / Standard Deviation, Higher Better)

```{r}
#nasdaq_metric_tbl <- read_rds(processed_data/nasdaq_metrics.rds)
```

```{r}
nasdaq_index_tbl <- read.csv("/Users/Admin/Desktop/FinalProject_Data607/nasdaq_index.csv")%>%
  clean_names()
```

```{r}
nasdaq_metric_screened_tbl <- stocks_metrics_query %>%
  
  inner_join(
    nasdaq_index_tbl %>% select(symbol, name, market_cap),
    by = "symbol"
  ) %>%
  
  filter(market_cap > 1e9) %>%
  
  arrange(-sd) %>%
  filter(
    sd < 1,
    count > 365 * 3,
    last_date == max(last_date)
  ) %>%
   mutate(reward_metric = 2500 * mean/sd) %>%
   mutate(desc = str_glue("
                        Symbol: {symbol}
                        mean: {round(mean, 3)}
                        SD: {round(sd, 3)}
                        N: {count}"))
```

# VISUALIZE SCREENING ---

```{r warning=FALSE}
g <- nasdaq_metric_screened_tbl %>%
    ggplot(aes(log(sd), log(mean)))+
    geom_point(aes(text = desc, color = reward_metric),
    alpha = 0.5, shape = 21, size = 4) + 
    geom_smooth() +
    scale_color_distiller(type = "div") +
    theme_minimal() +
    theme(
      panel.background = element_rect(fill = "black"),
      plot.background = element_rect(fill = "black"),
      text = element_text(colour = "white")
      ) +
  labs(title = "NASDAQ FINANCIAL ANALYSIS")

ggplotly(g)
```


# Shiny 

```{r}
# Define UI
ui <- fluidPage(
  titlePanel("NASDAQ FINANCIAL ANALYSIS"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("market_cap",
                  "Market Cap",
                  min = 1e9,
                  max = 1e12,
                  step = 1e9,
                  value = c(1e9, 1e12)),
      sliderInput("sd",
                  "Standard Deviation",
                  min = 0,
                  max = 1,
                  step = 0.01,
                  value = c(0, 1))
    ),
    mainPanel(
      plotlyOutput("plot")
    )
  )
)

# Define server
server <- function(input, output) {
  output$plot <- renderPlotly({
    filtered_tbl <- nasdaq_metric_screened_tbl %>%
      filter(
        market_cap >= input$market_cap[1] & market_cap <= input$market_cap[2],
        sd >= input$sd[1] & sd <= input$sd[2]
      )
    g <- filtered_tbl %>%
      ggplot(aes(log(sd), log(mean))) +
      geom_point(aes(text = desc, color = reward_metric),
                 alpha = 0.5, shape = 21, size = 4) + 
      geom_smooth() +
      scale_color_distiller(type = "div") +
      theme_minimal() +
      theme(
        panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill = "black"),
        text = element_text(colour = "white")
      ) +
      labs(title = "NASDAQ FINANCIAL ANALYSIS")
    ggplotly(g)
  })
}

# Run the app
shinyApp(ui = ui, server = server)
```




# Visualize best symbols ---

```{r}
n <- 9
best_symbol_tbl <- nasdaq_metric_screened_tbl %>%
      arrange(-reward_metric) %>%
      slice(1:n)
best_symbols <- best_symbol_tbl$symbol
stock_screen_data_tbl <- stocks_data_tbl %>%
  filter(symbol %in% best_symbols)
```

```{r, message=FALSE}
f <-stock_screen_data_tbl %>%
  left_join(
  best_symbol_tbl %>% select (symbol, name)
  ) %>%
  group_by(symbol, name) %>%
  plot_time_series(date, adjclose, .smooth = TRUE, .facet_ncol = 3, .interactive = F) +
  geom_line(color = "white") +
  theme_minimal() +
  theme(
    panel.background = element_rect(fill = "black"),
    plot.background = element_rect(fill = "black"),
    text = element_text(colour = "white"),
    strip.text = element_text(colour = "white")
  ) + 
  labs(title = "NASDAQ Financial Analysis")
ggplotly(f)
```


# Shiny for Both 


```{r}
# Define the UI
ui <- fluidPage(
  theme = shinytheme("slate"),
  titlePanel("NASDAQ Financial Analysis"),
  sidebarPanel(
    sliderInput(
      inputId = "market_cap",
      label = "Market Cap",
      min = 0,
      max = 1e12,
      value = c(1e9, 1e12),
      step = 1e8,
      sep = ""
    ),
    sliderInput(
      inputId = "sd",
      label = "Standard Deviation",
      min = 0,
      max = 1,
      value = c(0, 1),
      step = 0.01,
      sep = ""
    )
  ),
  mainPanel(
      plotlyOutput("plot1"),
      plotlyOutput("plot2")
  )
)

# Define the server
server <- function(input, output) {
  
  # Define the reactive filtered data
  filtered_data <- reactive({
    nasdaq_metric_screened_tbl <- stocks_metrics_query %>%
      inner_join(
        nasdaq_index_tbl %>% select(symbol, name, market_cap),
        by = "symbol"
      ) %>%
      filter(market_cap >= input$market_cap[1] & market_cap <= input$market_cap[2]) %>%
      arrange(-sd) %>%
      filter(
        sd >= input$sd[1],
        sd <= input$sd[2],
        count > 365 * 3,
        last_date == max(last_date)
      ) %>%
      mutate(reward_metric = 2500 * mean/sd) %>%
      mutate(desc = str_glue("
                        Symbol: {symbol}
                        mean: {round(mean, 3)}
                        SD: {round(sd, 3)}
                        N: {count}"))
    best_symbol_tbl <- nasdaq_metric_screened_tbl %>%
      arrange(-reward_metric) %>%
      slice(1:n)
    best_symbols <- best_symbol_tbl$symbol
    stocks_data_tbl %>%
      filter(symbol %in% best_symbols)
  })
  
  # Define the plot
  output$plot2 <- renderPlotly({
    filtered_data() %>%
      left_join(
        best_symbol_tbl %>% select(symbol, name)
      ) %>%
      group_by(symbol, name) %>%
      plot_time_series(date, adjclose, .smooth = TRUE, .facet_ncol = 3, .interactive = F) +
      geom_line(color = "white") +
      theme_minimal() +
      theme(
        panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill = "black"),
        text = element_text(colour = "white"),
        strip.text = element_text(colour = "white")
      ) +
      labs(title = "NASDAQ Financial Analysis")
    
  })
  
  # Plot the scatter plot of log(mean) vs. log(sd) with color-coded reward metric
  output$plot1 <- renderPlotly({
    filtered_tbl <- nasdaq_metric_screened_tbl %>%
      filter(
        market_cap >= input$market_cap[1] & market_cap <= input$market_cap[2],
        sd >= input$sd[1] & sd <= input$sd[2]
      )
    g <- filtered_tbl %>%
      ggplot(aes(log(sd), log(mean))) +
      geom_point(aes(text = desc, color = reward_metric),
                 alpha = 0.5, shape = 21, size = 4) + 
      geom_smooth() +
      scale_color_distiller(type = "div") +
      theme_minimal() +
      theme(
        panel.background = element_rect(fill = "black"),
        plot.background = element_rect(fill = "black"),
        text = element_text(colour = "white")
      ) +
      labs(title = "NASDAQ FINANCIAL ANALYSIS")
 
  })
}

# Run the app
shinyApp(ui = ui, server = server)
```



Conclusion: 


Source: 
https://therinspark.com/intro.html
source: https://www.nasdaq.com/market-activity/stocks/screener
https://www.kaggle.com/datasets/qks1lver/amex-nyse-nasdaq-stock-histories?select=fh_5yrs.csv
https://www.rdocumentation.org/packages/zoo/versions/1.8-12
https://search.r-project.org/CRAN/refmans/zoo/html/coredata.html
https://statisticsglobe.com/moving-average-maximum-median-sum-of-time-series-in-r
https://www.youtube.com/watch?v=rJawNrD3xlU&list=RDCMUCqQ_cxcNu1ekqHXDc-MbU_w&start_radio=1&t=3709s





